---
title: "Analysis"
author: "Anonymous"
date: "2024-05-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Introduction

This R Markdown file contains code for the analysis and plotting of data displayed in my dissertation. This file does NOT include code used to test assumptions of models.


## Preparing data

Setting the working directory as "Analysis" folder (**please adjust to suit your working directory**):

```{r}
setwd("~/1) MBiol Project/Green_Status_Analysis2/Analysis")
```


Loading in libraries, data, and functions:

```{r, warning = FALSE, message = FALSE}
source("Functions/Libraries.r")
source("Functions/Functions.r")

RawData <- read.csv("Data/Analysis_Data.csv")
```


Preparing data:

```{r}
#Function to prepare data: Adds new collumns with annual reproductive output (PARO),Species Recovery Scores as proportions, and Species Recovery Scores transformed to enable regular beta regression.
Data <- prepare_analysis_data(RawData)
```

Reviewing the data:

```{r}
headTail(Data) #shows first and last rows
```

## Analysis


## 1. Characterising Recovery Status between and across families


Mean Species Recovery Score for all species:
```{r}
mean(Data$SRS01)
```

Standard Deviation: 
```{r}
SRSsd <- sd(Data$SRS01)
SRSsd
```

Standard Error:
```{r}
(SRSsd/sqrt(20))
```

Mean Species Recovery Score for each Family:
```{r}
Fammean <- tapply(Data$SRS01,Data$Family,mean)
Fammean
```

Standard deviation for each Family:
```{r}
Famsd <- tapply(Data$SRS01,Data$Family,sd)
Famsd
```

Standard error for each Family:
```{r}
Famse <- (Famsd/sqrt(20))
Famse
```

95% Confidence interval: Upper bound
```{r}
Famup <- Fammean + (1.96*Famse)
Famup
```

95% Confidence interval: Lower bound
```{r}
Famlo <- Fammean - (1.96*Famse)
Famlo
```


Plotting mean Species Recovery Score:

```{r}
DataSE <- summarySE(Data, measurevar="SRS01", groupvars=c("Family"))

# Grouped by family
tiff(filename = "Figures/SRSbyfam.tiff", width = 900, height = 900, units = "px") #saves figure below as tiff
set.seed(104)
ggplot(DataSE, aes(x=Family, y=SRS01, colour = Family, fill = Family)) + 
  geom_hline(yintercept = 0.001, linetype = "dashed", color = "grey", size = 1) +
  geom_hline(yintercept = 0.2, linetype = "dashed", color = "grey", size = 1) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "grey", size = 1) +
  geom_hline(yintercept = 0.8, linetype = "dashed", color = "grey", size = 1) +
  geom_hline(yintercept = 0.999, linetype = "dashed", color = "grey", size = 1) +
    geom_point(stat="identity", size = 9, show.legend = FALSE) +
    geom_errorbar(aes(ymin=SRS01-ci, ymax=SRS01+ci),
                  width=.3,                    # Width of the error bars
                  position=position_dodge(.9), size = 2, show.legend = FALSE) +
  geom_quasirandom(data = Data, width = .2, alpha = 0.5, height = 0, size = 4.5, show.legend = FALSE) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
	theme(text = element_text(size = 30))+
  theme(axis.title = element_text(size=35, face = "bold"), axis.title.y = element_text(margin=margin(r=30))) +
  scale_colour_manual(values=c("#24B14D", "#FFC107")) +
  ylab("Mean Species Recovery Score") +
  ylim(0,1)
dev.off() #Final code to save as tiff
```


Plotting Mean Upper and Lower bounds of Species Recovery Score:

```{r}
# Upper and Lower Estimates

DataUSE <- summarySE(Data, measurevar="U01", groupvars=c("Family"))

DataLSE <- summarySE(Data, measurevar="L01", groupvars=c("Family"))


# Upper
tiff(filename = "Figures/UpperSRSbyfam.tiff", width = 900, height = 900, units = "px") #saves figure below as tiff
set.seed(104)
ggplot(DataUSE, aes(x=Family, y=U01, colour = Family, fill = Family)) + 
  geom_hline(yintercept = 0.001, linetype = "dashed", color = "grey", size = 1) +
  geom_hline(yintercept = 0.2, linetype = "dashed", color = "grey", size = 1) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "grey", size = 1) +
  geom_hline(yintercept = 0.8, linetype = "dashed", color = "grey", size = 1) +
  geom_hline(yintercept = 0.999, linetype = "dashed", color = "grey", size = 1) +
    geom_point(stat="identity", size = 9, show.legend = FALSE) +
    geom_errorbar(aes(ymin=U01-ci, ymax=U01+ci),
                  width=.3,                    # Width of the error bars
                  position=position_dodge(.9), size = 2, show.legend = FALSE) +
  geom_quasirandom(data = Data, width = .2, alpha = 0.5, height = 0, size = 4.5, show.legend = FALSE) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
	theme(text = element_text(size = 30))+
  theme(axis.title = element_text(size=35, face = "bold"), axis.title.y = element_text(margin=margin(r=30))) +
  scale_colour_manual(values=c("#24B14D", "#FFC107")) +
  ylab("Mean Upper Bound of Species Recovery Score") +
  ylim(0,1)
dev.off() #Final code to save as tiff


# Lower
tiff(filename = "Figures/LowerSRSbyfam.tiff", width = 900, height = 900, units = "px") #saves figure below as tiff
set.seed(104)
ggplot(DataLSE, aes(x=Family, y=L01, colour = Family, fill = Family)) + 
  geom_hline(yintercept = 0.001, linetype = "dashed", color = "grey", size = 1) +
  geom_hline(yintercept = 0.2, linetype = "dashed", color = "grey", size = 1) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "grey", size = 1) +
  geom_hline(yintercept = 0.8, linetype = "dashed", color = "grey", size = 1) +
  geom_hline(yintercept = 0.999, linetype = "dashed", color = "grey", size = 1) +
    geom_point(stat="identity", size = 9, show.legend = FALSE) +
    geom_errorbar(aes(ymin=L01-ci, ymax=L01+ci),
                  width=.3,                    # Width of the error bars
                  position=position_dodge(.9), size = 2, show.legend = FALSE) +
  geom_quasirandom(data = Data, width = .2, alpha = 0.5, height = 0, size = 4.5, show.legend = FALSE) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
	theme(text = element_text(size = 30))+
  theme(axis.title = element_text(size=35, face = "bold"), axis.title.y = element_text(margin=margin(r=30))) +
  scale_colour_manual(values=c("#24B14D", "#FFC107")) +
  ylab("Mean Lower Bound of Species Recovery Score") +
  ylim(0,1)
dev.off() #Final code to save as tiff
```

## 2. Testing for predictors of Recovery Status

Performing beta regression model

```{r}
#model containing Family, Annual Reproductive Output, Body Mass Class, and all of their interactions
Data$BMC <- factor(Data$BMC, levels = c("S", "M", "L"))
BMCmodel = betareg(Data$best_trans ~ Family*BMC  + Family*PARO + PARO*BMC, data = Data)
```

Displaying outputs of the model

```{r}
summary(BMCmodel)
```

Perfoming beta regression without potential reproductive output

```{r}
Data$BMC <- factor(Data$BMC, levels = c("S", "M", "L"))
BMCmodel2 = betareg(Data$best_trans ~ Family*BMC, data = Data)
joint_tests(BMCmodel2)
summary(BMCmodel2)
```

Calculating Estimated Marginal Means

```{r}
marginalFB = emmeans(BMCmodel2, ~ BMC*Family) #Calculating estimated marginal means
```


Plotting Estimated Marginal Means
```{r}
SumFB <- SumFunction(marginalFB) #Prepping EMMs for plotting

SumFB$BMC <- factor(SumFB$BMC, levels = c("S", "M", "L"))

tiff(filename = "Figures/BMCFamily estimated means SRS.tiff", width = 900, height = 900, units = "px") #saves figure below as tiff
set.seed(50)
	ggplot(SumFB,                ### The data frame to use.
       aes(x = BMC,
           y = emmean, colour = Family)) +
	geom_hline(yintercept = 0.001, linetype = "dashed", color = "grey", size = 1) +
  geom_hline(yintercept = 0.2, linetype = "dashed", color = "grey", size = 1) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "grey", size = 1) +
  geom_hline(yintercept = 0.8, linetype = "dashed", color = "grey", size = 1) +
  geom_hline(yintercept = 0.999, linetype = "dashed", color = "grey", size = 1) +
  	 geom_errorbar(aes(ymin = asymp.LCL,
                     ymax = asymp.UCL),
                     width = 0.5,
                     size  = 1.5, position = "dodge") +
  	 geom_point(shape = 15,
              size  = 10, position = position_dodge(0.5)) +
	   geom_quasirandom(data = Data, width = .15, alpha = 0.5, size = 4, show.legend = FALSE,  dodge.width = 0.5, aes(x = BMC, y = SRS01, colour = Family)) +
   	 theme(axis.title   = element_text(face  = "bold")) +
	    theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
	theme(text = element_text(size = 35),  plot.margin = margin(1, 1, 1, 1, "cm"), legend.title = element_text(size=25, face="bold"), legend.text = element_text(size = 20))+
  theme(axis.title = element_text(size=30, face = "bold"), axis.title.y = element_text(margin=margin(r=30)), axis.title.x = element_text(vjust=0), margin=margin(r=40)) +
  scale_colour_manual(values=c("#24B14D", "#FFC107")) +
	  scale_x_discrete(labels=c("Small", "Medium", "Large"))+
    	 xlab("Body Mass Class") +
	 ylab("Estimated Marginal Mean Species Recovery Score") +
	 ylim(0,1.05) + 
annotate("text",x = c(1.9,0.9,1.1,2.1,3.05,2.8), y = SumFB$asymp.UCL + 0.026, label = gsub("", "", SumFB$.group), size = 10)


dev.off() #final command to save as .tiff
```


Repeating with Upper and Lower bounds of Species Recovery Score

Upper Bound

```{r}
#Upper bound model
Umodel = betareg(Data$upper_trans ~ Family*BMC, data = Data) 
joint_tests(Umodel)
summary(Umodel)
```


Lower Bound

```{r}
#Lower bound model
Lmodel = betareg(Data$lower_trans ~ Family*BMC, data = Data)
joint_tests(Lmodel)
summary(Lmodel)
```


Calculating Estimated Marginal Means
```{r}
marginalU = emmeans(Umodel, ~ BMC*Family) #Calculating estimated marginal means

SumU <- SumFunction(marginalU) #Prepping EMMs for plotting


marginalL = emmeans(Lmodel, ~ BMC*Family) #Calculating estimated marginal means

SumL <- SumFunction(marginalL) #Prepping EMMs for plotting
```


Plotting Estimated Marginal Means
```{r}
SumU$BMC <- factor(SumU$BMC, levels = c("S", "M", "L"))

tiff(filename = "Figures/Upper estimated means SRS.tiff", width = 900, height = 900, units = "px") #saves figure below as tiff
set.seed(60)
	ggplot(SumU,                ### The data frame to use.
       aes(x = BMC,
           y = emmean, colour = Family)) +
	geom_hline(yintercept = 0.001, linetype = "dashed", color = "grey", size = 1) +
  geom_hline(yintercept = 0.2, linetype = "dashed", color = "grey", size = 1) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "grey", size = 1) +
  geom_hline(yintercept = 0.8, linetype = "dashed", color = "grey", size = 1) +
  geom_hline(yintercept = 0.999, linetype = "dashed", color = "grey", size = 1) +
  	 geom_errorbar(aes(ymin = asymp.LCL,
                     ymax = asymp.UCL),
                     width = 0.5,
                     size  = 1.5, position = "dodge") +
  	 geom_point(shape = 15,
              size  = 10, position = position_dodge(0.5)) +
	   geom_quasirandom(data = Data, width = .1, alpha = 0.5, size = 4, show.legend = FALSE,  dodge.width = 0.5, aes(x = BMC, y = U01, colour = Family)) +
   	 theme(axis.title   = element_text(face  = "bold")) +
	    theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
	theme(text = element_text(size = 35),  plot.margin = margin(1, 1, 1, 1, "cm"), legend.title = element_text(size=25, face="bold"), legend.text = element_text(size = 20))+
  theme(axis.title = element_text(size=30, face = "bold"), axis.title.y = element_text(margin=margin(r=30)), axis.title.x = element_text(vjust=0), margin=margin(r=40)) +
  scale_colour_manual(values=c("#24B14D", "#FFC107")) +
	  scale_x_discrete(labels=c("Small", "Medium", "Large"))+
    	 xlab("Body Mass Class") +
	 ylab("Estimated Marginal Mean \n Upper Bound of Species Recovery Score") +
	 ylim(0,1.05) + 
annotate("text",x = c(1.9, 0.9, 1.15, 2.1, 3.1, 2.8), y = SumU$asymp.UCL + 0.026, label = gsub("", "", SumU$.group), size = 10)


dev.off() #final command to save as .tiff

# LOWER

SumL$BMC <- factor(SumL$BMC, levels = c("S", "M", "L"))

tiff(filename = "Figures/Lower estimated means SRS.tiff", width = 900, height = 900, units = "px") #saves figure below as tiff
set.seed(60)
	ggplot(SumL,                ### The data frame to use.
       aes(x = BMC,
           y = emmean, colour = Family)) +
	geom_hline(yintercept = 0.001, linetype = "dashed", color = "grey", size = 1) +
  geom_hline(yintercept = 0.2, linetype = "dashed", color = "grey", size = 1) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "grey", size = 1) +
  geom_hline(yintercept = 0.8, linetype = "dashed", color = "grey", size = 1) +
  geom_hline(yintercept = 0.999, linetype = "dashed", color = "grey", size = 1) +
  	 geom_errorbar(aes(ymin = asymp.LCL,
                     ymax = asymp.UCL),
                     width = 0.5,
                     size  = 1.5, position = "dodge") +
  	 geom_point(shape = 15,
              size  = 10, position = position_dodge(0.5)) +
	   geom_quasirandom(data = Data, width = .15, alpha = 0.5, size = 4, show.legend = FALSE,  dodge.width = 0.5, aes(x = BMC, y = L01, colour = Family)) +
   	 theme(axis.title   = element_text(face  = "bold")) +
	    theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
	theme(text = element_text(size = 35),  plot.margin = margin(1, 1, 1, 1, "cm"), legend.title = element_text(size=25, face="bold"), legend.text = element_text(size = 20))+
  theme(axis.title = element_text(size=30, face = "bold"), axis.title.y = element_text(margin=margin(r=30)), axis.title.x = element_text(vjust=0), margin=margin(r=40)) +
  scale_colour_manual(values=c("#24B14D", "#FFC107")) +
	  scale_x_discrete(labels=c("Small", "Medium", "Large"))+
    	 xlab("Body Mass Class") +
	 ylab("Estimated Marginal Mean \n Lower Bound of Species Recovery Score") +
	 ylim(0,1.05) + 
annotate("text",x = c(0.9, 1.9, 2.1, 3.05, 1.05, 2.8), y = SumL$asymp.UCL + 0.026, label = gsub("", "", SumL$.group), size = 10)

dev.off() #final command to save as .tiff
```

## 3. Testing for differences between predictors of Recovery Status and Extinction Risk

Performing ordered logistic regression to test for predictors of Red List Category

Body Mass Class and Family

```{r}
#model containing Family, Body Mass Class, and Potential Reproductive Output
Data$Red_List_Category <- factor(Data$Red_List_Category, levels = c("Critically Endangered", "Endangered", "Vulnerable", "Near Threatened", "Least Concern"))
Data$BMC <- factor(Data$BMC, levels = c("S", "M", "L"))

m <- polr(Red_List_Category ~ BMC * Family, data = Data%>%  filter(!row_number() %in% c(5)), Hess=TRUE)
joint_tests(m)
summary(m)
```

Displaying outputs and calculating p values

```{r}
## store table
ctable <- coef(summary(m))

## calculate and store p values
p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2

## combined table
(ctable <- cbind(ctable, "p value" = p))
```


```{r}
#model containing Body Mass Class
m2 <- polr(Red_List_Category ~ BMC, data = Data%>%  filter(!row_number() %in% c(5)), Hess=TRUE)
joint_tests(m2)
summary(m2)
```


```{r}
## store table
ctable2 <- coef(summary(m2))

## calculate and store p values
p2 <- pnorm(abs(ctable2[, "t value"]), lower.tail = FALSE) * 2

## combined table
(ctable2 <- cbind(ctable2, "p value" = p2))
```


Annual Reproductive Ouput and Family

```{r}
#model containing Family, Body Mass Class, and Potential Reproductive Output
Data$Red_List_Category <- factor(Data$Red_List_Category, levels = c("Critically Endangered", "Endangered", "Vulnerable", "Near Threatened", "Least Concern"))
Data$BMC <- factor(Data$BMC, levels = c("S", "M", "L"))

m3 <- polr(Red_List_Category ~ PARO * Family, data = Data%>%  filter(!row_number() %in% c(5)), Hess=TRUE)
joint_tests(m3)
summary(m3)
```

Displaying outputs and calculating p values

```{r}
## store table
ctable3 <- coef(summary(m3))

## calculate and store p values
p3 <- pnorm(abs(ctable3[, "t value"]), lower.tail = FALSE) * 2

## combined table
(ctable3 <- cbind(ctable3, "p value" = p3))
```


## 4. Testing for Predictors of Uncertainty in Recovery Status

Performing linear regression

```{r}
Data$BMC <- factor(Data$BMC, levels = c("S", "M", "L"))
Data$Red_List_Category <- factor(Data$Red_List_Category, levels = c("Critically Endangered", "Endangered", "Vulnerable", "Near Threatened", "Least Concern"))
uncertmodel <- lm(Range ~ BMC*Family + Red_List_Category, data = Data%>%  filter(!row_number() %in% c(5)))
joint_tests(uncertmodel)
summary(uncertmodel)
```

Calculating Estimated Marginal Means for Red List Category

```{r}
marginalUR = emmeans(uncertmodel, ~ Red_List_Category) #Calculating estimated marginal means

SumUR <- SumFunction(marginalUR) #Prepping EMMs for plotting
```

Plotting Estimated Marginal Means

```{r}
tiff(filename = "Figures/UncertRed estimated means SRs.tiff", width = 650, height = 900, units = "px") #saves figure below as tiff

	ggplot(SumUR,                ### The data frame to use.
       aes(x = Red_List_Category,
           y = emmean)) +
	  geom_hline(yintercept=0, linetype="dashed", colour = "darkgray", size = 1.5) +
	  #geom_quasirandom(data = Data, width = .2, alpha = 0.8, size = 4, colour = "grey", show.legend = FALSE,  dodge.width = 0.5, aes(x = Red_List_Category, y = Range)) +
  	 geom_errorbar(aes(ymin = lower.CL,
                     ymax = upper.CL),
                     width = 0.5,
                     size  = 1.2, position = "dodge") +
  	 geom_point(shape = 15,
              size  = 10, position = position_dodge(0.5)) +
   	 theme(axis.title   = element_text(face  = "bold")) +
	    theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
	theme(text = element_text(size = 35),  plot.margin = margin(1, 1, 1, 1, "cm"), legend.title = element_text(size=25, face="bold"), legend.text = element_text(size = 20))+
  theme(axis.title = element_text(size=30, face = "bold"), axis.title.y = element_text(margin=margin(r=30)), axis.title.x = element_text(vjust=0), margin=margin(r=40)) +
    	 xlab("Red List Category") +
	 ylab("Uncertainty in Species Recovery Score (%)") +
	 ylim(-20,40) + 
	 scale_x_discrete(limits = rev(levels(SumUR$Red_List_Category)), labels=c("LC", "NT", "VU", "EN")) +
annotate("text",x = SumUR$Red_List_Category, y = SumUR$upper.CL + 2, label = gsub("", "", SumUR$.group), size = 10)


dev.off() #final command to save as .tiff
```

Calculating Estimated Marginal Means of Body Mass Class by Family

```{r}
marginalUBF = emmeans(uncertmodel, ~ BMC*Family) #Calculating estimated marginal means

SumUBF <- SumFunction(marginalUBF) #Prepping EMMs for plotting
```

Plotting Estimated Marginal Means 

```{r}
SumUBF$BMC <- factor(SumUBF$BMC, levels = c("S", "M", "L"))

tiff(filename = "Figures/UncertFamBod estimated means SRs.tiff", width = 650, height = 900, units = "px") #saves figure below as tiff

	ggplot(SumUBF,                ### The data frame to use.
       aes(x = BMC,
           y = emmean, colour = Family)) +
	  geom_hline(yintercept=0, linetype="dashed", color = "grey", size = 1.5) +
  	 geom_errorbar(aes(ymin = lower.CL,
                     ymax = upper.CL),
                     width = 0.5,
                     size  = 1.2, position = position_dodge(width = 0.7)) +
  	 geom_point(shape = 15,
              size  = 10, position = position_dodge(0.7)) +
   	 theme(axis.title   = element_text(face  = "bold")) +
	    theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
	theme(text = element_text(size = 35),  plot.margin = margin(1, 1, 1, 1, "cm"), legend.title = element_text(size=25, face="bold"), legend.text = element_text(size = 20))+
  theme(axis.title = element_text(size=30, face = "bold"), axis.title.y = element_text(margin=margin(r=30)), axis.title.x = element_text(vjust=0), margin=margin(r=40)) +
	  scale_colour_manual(values=c("#24B14D", "#FFC107")) +
	  scale_x_discrete(labels=c("Small", "Medium", "Large"))+
    	 xlab("Body Mass Class") +
	 ylab("Uncertainty in Species Recovery Score (%)") +
	 ylim(-20,40) + 
annotate("text",x = c(1.2,1.8,0.8,3.15,2.77,2.1), y = SumUBF$upper.CL + 2, label = gsub("", "", SumUBF$.group), size = 10)


dev.off() #final command to save as .tiff
```
